\documentclass[12pt]{article}

% packages
\usepackage{kantlipsum}
\usepackage[margin=1in]{geometry}
\usepackage[labelfont=it]{caption}
\usepackage[table]{xcolor}
\usepackage{subcaption,framed,colortbl,multirow}
\usepackage{amsmath,amsthm,amssymb,wasysym,mathrsfs,mathtools}
\usepackage{tikz,graphicx,pgf,pgfplots}
\usetikzlibrary{arrows, angles, quotes, decorations.pathreplacing, math, patterns, calc}
\pgfplotsset{compat=1.16}

% custom commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\p}{^{\prime}}
\newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\setlength{\fboxsep}{4pt}
\newcommand{\generic}[2]{\section*{#1}\begin{center}\framebox{\begin{minipage}{\textwidth-10pt}#2\end{minipage}}\end{center}}
\newcommand{\ex}[2]{\generic{Exercise #1}{#2}}
\newcommand{\prob}[2]{\generic{Problem #1}{#2}}
\newcommand{\ques}[2]{\generic{Question #1}{#2}}

\newenvironment{drawing}{\begin{center}\begin{tikzpicture}}{\end{tikzpicture}\end{center}}

 
\begin{document}
 
\title{Assignment 0\\
    \large MATH CS 117 Intro to Real Analysis}
\author{Harry Coleman}
\date{April 2, 2020}
\maketitle

\prob{1}{
    Relative to California, in what time zone will you be living during the Spring Quarter? Please give your answer in the form $+x$ if you are $x$ hours ahead and $-x$ if you are $x$ hours behind.
}

California Time $+0$.

\prob{2}{
    True or false? $0 \in \N$.
}

False. (For this class, based on the notes. But depending on the context, some people include 0 in the natural numbers.)

\prob{3}{
    True or false? Every set in $\R$ is bounded above.
}

False. 

\newpage
\prob{4}{
    True or false? $\Q$ is an ordered field.
}

True, but $\Q$ is not a \emph{complete} ordered field since it does not satisfy the least upper bound property.

\prob{5}{
    True or false? $\R$ satisfies the least upper bound principle.
}

True.


\end{document}